---
title: "Short-Term Rentals (STR) in Cambridge, MA"
output: 
  flexdashboard::flex_dashboard:
    theme: spacelab
    source_code: embed
    orientation: rows
runtime: shiny
---

```{r setup, include=FALSE}
library(plyr)
library(flexdashboard)
library(shinyWidgets)
library(shiny)
library(tidyverse)
library(janitor)
library(magrittr)
library(leaflet)
library(DT)
library(plotly)
library(shinydashboard)
```

```{r global, include = FALSE}
url = "https://data.cambridgema.gov/resource/wxgv-w968.csv?$query=SELECT%0A%20%20%60recordid%60%2C%0A%20%20%60issue_date%60%2C%0A%20%20%60status%60%2C%0A%20%20%60location%60%2C%0A%20%20%60latitude%60%2C%0A%20%20%60longitude%60%2C%0A%20%20%60please_select_the_appropriate_str_type%60%2C%0A%20%20%60type_of_property%60%2C%0A%20%20%60please_explain_below%60%2C%0A%20%20%60is_the_short_term_rental_unit_part_of_a_condominium_association%60%2C%0A%20%20%60total_number_of_legal_bedrooms_in_the_short_term_rental_unit%60%2C%0A%20%20%60number_of_legal_bedrooms_to_be_rented_as_part_of_this_application%60%2C%0A%20%20%60maximum_number_or_people_to_which_the_short_term_rental_will_be_rented%60%2C%0A%20%20%60will_the_kitchen_be_available_to_the_renter%60%2C%0A%20%20%60number_of_bathrooms_available_to_the_renter%60%2C%0A%20%20%60airbnb%60%2C%0A%20%20%60homeaway%60%2C%0A%20%20%60flipkey%60%2C%0A%20%20%60vrbo%60%2C%0A%20%20%60craigslist%60%2C%0A%20%20%60couch_surfing%60%2C%0A%20%20%60boston_rentals%60%2C%0A%20%20%60other%60%2C%0A%20%20%60please_list_additional_hosting_platforms%60%2C%0A%20%20%60all_rental_services%60%2C%0A%20%20%60%3A%40computed_region_guic_hr4a%60%2C%0A%20%20%60%3A%40computed_region_v7jj_366k%60%2C%0A%20%20%60%3A%40computed_region_e4yd_rwk4%60%2C%0A%20%20%60%3A%40computed_region_swkg_bavi%60%2C%0A%20%20%60%3A%40computed_region_rffn_qbt6%60%2C%0A%20%20%60%3A%40computed_region_rcj3_ccgu%60"
data <- read_csv(url)
colnames <- c("id", "issue_date", "status", "location", "lat", "lon", "str_type", "property_type", "property_type_desc", "condo", "total_bedrooms", "rented_bedrooms", "max_renter_cap", "kitchen", "bathrooms", "airbnb", "homeaway", "flipkey", "vrbo", "craigslist", "couch_surfing", "boston_rentals", "other", "other_plat_desc", "all_rental_serv", "police_nei_regions", "police_res_districts", "census_blocks_2010", "cdd_zoning", "neighborhoods", "zipcode")


property_type_from <- c("Single-family detached dwelling","Two-family dwelling","Three-family structure", "Multi-family dwelling of four units or less, all of which are owned by the operator", "Other")
property_type_to <- c("Single-family","Two-family","Three-family", "Multi-family", "Other")

df <- data %>%
  set_colnames(colnames) %>%
  mutate(issue_date = as.Date(issue_date),
         str_type = str_to_sentence(gsub("^.*an\\s+(.*)\\s+short-term.*$", "\\1", str_type)),
         property_type = factor(mapvalues(property_type, from=property_type_from, to=property_type_to), levels = property_type_to),
         address = gsub("Cambridge,.*$", "", location),
         zipcode = if_else(str_length(gsub("^.*MA\\s+(\\w+).*", "\\1", location)) == 5, gsub("^.*MA\\s+(\\w+).*", "\\1", location), "02139"),
         airbnb = ifelse(is.na(airbnb), FALSE, airbnb),
         homeaway = ifelse(is.na(homeaway), FALSE, homeaway),
         flipkey = ifelse(is.na(flipkey), FALSE, flipkey),
         vrbo = ifelse(is.na(vrbo), FALSE, vrbo),
         craigslist = ifelse(is.na(craigslist), FALSE, craigslist),
         couch_surfing = ifelse(is.na(couch_surfing), FALSE, couch_surfing),
         boston_rentals = ifelse(is.na(boston_rentals), FALSE, boston_rentals)
         ) %>%
  select(-status, -all_rental_serv, -property_type_desc, -other, -other_plat_desc, -neighborhoods, -police_nei_regions, -police_res_districts, -census_blocks_2010, -cdd_zoning, -location) %>%
  clean_names()


zipcode_list <- sort(unique(df$zipcode))
str_type_list <- sort(unique(df$str_type))
property_type_list <- sort(unique(df$property_type))
total_bedrooms_list <- sort(unique(df$total_bedrooms))
rented_bedrooms_list <- sort(unique(df$rented_bedrooms))
max_renter_cap_list <- sort(unique(df$max_renter_cap))
bathrooms_list <- sort(unique(df$bathrooms))
```

# Spatial distribution {data-icon="ion-earth"}

## Inputs {.sidebar data-width="300"}

```{r}
pickerInput("zipcode",
            label = "Select zipcodes",
            choices = zipcode_list,
            selected = zipcode_list, 
            options = list(`actions-box` = TRUE),
          multiple = TRUE)

pickerInput("str_type",
            label = "Select Short-Term Rental types",
            choices = str_type_list,
            selected = str_type_list[1], 
            options = list(`actions-box` = TRUE),
          multiple = TRUE)

pickerInput("property_type",
            label = "Select property types",
            choices = property_type_to,
            selected = property_type_to[1], 
            options = list(`actions-box` = TRUE),
          multiple = TRUE)

#sliderInput("total_bedrooms", "Total Bedrooms",
#                  min = min(total_bedrooms_list), 
#                  max = max(total_bedrooms_list),
#                  value = c(min(total_bedrooms_list),max(total_bedrooms_list)),
#                  sep = "")

sliderInput("rented_bedrooms", "Select number of rented bedrooms",
                  min = min(rented_bedrooms_list), 
                  max = max(rented_bedrooms_list),
                  value = c(min(rented_bedrooms_list),max(rented_bedrooms_list)),
                  sep = "")

sliderInput("max_renter_cap", "Select maximum renter capacity",
                  min = min(max_renter_cap_list), 
                  max = max(max_renter_cap_list),
                  value = c(min(max_renter_cap_list),max(max_renter_cap_list)),
                  sep = "")

sliderInput("bathrooms", "Select number of bathrooms",
                  min = min(bathrooms_list), 
                  max = max(bathrooms_list),
                  value = c(min(bathrooms_list),max(bathrooms_list)),
                  sep = "")

#pickerInput("kitchen",
#            label = "Kitchen available to the renter",
#            choices = kitchen_list,
#            selected = kitchen_list, 
#          multiple = TRUE)

#pickerInput("condo",
#            label = "Part of a condominium association",
#            choices = condo_list,
#            selected = condo_list, 
#          multiple = TRUE)


h5("Listed on any of these services")
checkboxInput("airbnb", "Airbnb", value = TRUE)
checkboxInput("homeaway", "HomeAway", value = TRUE)
checkboxInput("flipkey", "FlipKey", value = TRUE)
checkboxInput("vrbo", "VRBO", value = TRUE)
checkboxInput("craigslist", "Craigslist", value = TRUE)
#checkboxInput("couch_surfing", "Couch Surfing", value = TRUE)
#checkboxInput("boston_rentals", "Boston Rentals", value = TRUE)

actionButton("update1", "Update")
```

## Column

### Map of selected Short-Term Rentals

```{r}
content <- paste("<b>","ID:","</b>",df$id,"</br>",
                 "<b>","Address:","</b>",df$address,"</br>",
                 "<b>","Zipcode:","</b>",df$zipcode,"</br>",
                 "------------------------------------------------------","</br>",
                 "<b>","STR type:","</b>",df$str_type,"</br>",
                 "<b>","Property type:","</b>",df$property_type,"</br>",
                 "<b>","Total bedrooms:","</b>",df$total_bedrooms,"</br>",
                 "<b>","Rented bedrooms:","</b>",df$rented_bedrooms,"</br>",
                 "<b>","Max renter capacity:","</b>",df$max_renter_cap,"</br>",
                 "<b>","Bathrooms:","</b>", df$bathrooms,"</br>",
                 "<b>","Kitchen:","</b>", df$kitchen,"</br>",
                 "<b>","Part of condominium association:","</b>", df$condo,"</br>",
                 "<b>","STR certificate issue date:","</b>", df$issue_date
                 )



df_map1 = eventReactive(input$update1, {
    
    df_blank = df %>% filter(id=="dummy")
    df_airbnb =   df_homeaway = df_flipkey = df_vrbo = df_craiglist = NULL
    if (input$airbnb == TRUE) {df_airbnb <- df %>% filter(airbnb==TRUE)}
    if (input$homeaway == TRUE) {df_homeaway <- df %>% filter(homeaway==TRUE)}
    if (input$flipkey == TRUE) {df_flipkey <- df %>% filter(flipkey==TRUE)}
    if (input$vrbo == TRUE) {df_vrbo <- df %>% filter(vrbo==TRUE)}
    if (input$craigslist == TRUE) {df_craiglist <- df %>% filter(craigslist==TRUE)}

    df <- unique(bind_rows(df_blank, df_airbnb, df_homeaway, df_flipkey, df_vrbo, df_craiglist)) %>%
      filter(zipcode %in% input$zipcode,
             str_type %in% input$str_type,
             property_type %in% input$property_type,
             between(rented_bedrooms, input$rented_bedrooms[1], input$rented_bedrooms[2]),
             between(max_renter_cap, input$max_renter_cap[1], input$max_renter_cap[2]),
             between(bathrooms, input$bathrooms[1], input$bathrooms[2]))
  })

greenLeafIcon <- makeIcon(
  iconUrl = "588a667fd06f6719692a2d19.png",
  iconWidth = 15, iconHeight = 15)

renderLeaflet({
leaflet(options =leafletOptions(minZoom =11,maxZoom =16)) %>%
  setView(lng = -71.1, lat =42.38, zoom =12) %>%
  addTiles() %>%
  addMarkers(lng = ~lon,
             lat = ~lat,
             data = df_map1(),
             popup = content, 
             icon = greenLeafIcon) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMiniMap()
})

```

## Column

### Information of selected Short-Term Rentals

```{r}

table_colnames <- c("Address", "Zipcode", "STR Type", "Property Type", "Total Bedrooms", "Rented Bedrooms", "Max. Renter Capacity", "Bathrooms", "Kitchen", "Condominium Association")

df_table1 <- eventReactive(input$update1, {
    df_blank = df %>% filter(id=="dummy")
    df_airbnb =   df_homeaway = df_flipkey = df_vrbo = df_craiglist = NULL
    if (input$airbnb == TRUE) {df_airbnb <- df %>% filter(airbnb==TRUE)}
    if (input$homeaway == TRUE) {df_homeaway <- df %>% filter(homeaway==TRUE)}
    if (input$flipkey == TRUE) {df_flipkey <- df %>% filter(flipkey==TRUE)}
    if (input$vrbo == TRUE) {df_vrbo <- df %>% filter(vrbo==TRUE)}
    if (input$craigslist == TRUE) {df_craiglist <- df %>% filter(craigslist==TRUE)}

    df <- unique(bind_rows(df_blank, df_airbnb, df_homeaway, df_flipkey, df_vrbo, df_craiglist)) %>%
      filter(zipcode %in% input$zipcode,
             str_type %in% input$str_type,
             property_type %in% input$property_type,
             between(rented_bedrooms, input$rented_bedrooms[1], input$rented_bedrooms[2]),
             between(max_renter_cap, input$max_renter_cap[1], input$max_renter_cap[2]),
             between(bathrooms, input$bathrooms[1], input$bathrooms[2])) %>%
      select(address, zipcode, str_type, property_type, total_bedrooms, rented_bedrooms, max_renter_cap, bathrooms, kitchen, condo) %>%
    set_colnames(table_colnames)
  })
  
renderDT({
    datatable(df_table1(), 
              options = list(paging = TRUE,
                             searching = TRUE,
                             orderClasses = TRUE))
  })

```

# Statistical analysis {data-icon="ion-stats-bars"}

## Inputs {.sidebar data-width="300"}

```{r}
quant_vars <- c("total_bedrooms", "rented_bedrooms", "max_renter_cap", "bathrooms")
qual_vars <- c("zipcode", "str_type", "property_type", "kitchen", "condo", "airbnb", "flipkey", "vrbo", "craigslist")

h5(HTML("<b>Parameters for boxplot</b>"))

pickerInput("y",
             label = "Select variable of interest",
             choices = quant_vars,
             selected = quant_vars[1], 
             multiple = FALSE)

pickerInput("x",
            label = "Select grouping variable",
             choices = qual_vars,
             selected = qual_vars[1], 
            multiple = FALSE)


h5(HTML("<b>Filters</b>"))

pickerInput("zipcode2",
            label = "Select zipcodes",
            choices = zipcode_list,
            selected = zipcode_list, 
            options = list(`actions-box` = TRUE),
          multiple = TRUE)

pickerInput("str_type2",
            label = "Select Short-Term Rental types",
            choices = str_type_list,
            selected = str_type_list[1], 
            options = list(`actions-box` = TRUE),
          multiple = TRUE)

pickerInput("property_type2",
            label = "Select property types",
            choices = property_type_to,
            selected = property_type_to[1], 
            options = list(`actions-box` = TRUE),
          multiple = TRUE)

actionButton("update2", "Update")
```

## Column

### Boxplot analysis

```{r}
#df[df == "TRUE"] <- "Yes"
#df[toString(df)== "TRUE"] = "Yes"
#print(df$airbnb)
#replace(df$airbnb, toString(df$airbnb) == "FALSE", "No") 
#mapping <-list("Total Bedrooms" = "total_bedrooms"")
df_boxplot = eventReactive(input$update2, {
  df %>% 
    filter(zipcode %in% input$zipcode2,
             str_type %in% input$str_type2,
             property_type %in% input$property_type2) %>%
    mutate(x = .data[[input$x]],
           y = .data[[input$y]])
})

x <- eventReactive(input$update2, input$x)
y <- eventReactive(input$update2, input$y)

renderPlotly({
p1 <- ggplotly(
  ggplot(data = df_boxplot(), mapping = aes(x = x, y = y)) +
  geom_boxplot(aes(fill = x)) +  
  geom_jitter(width=0.15, height=0.1, alpha = 0.5) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  theme_bw() +
  theme(legend.position="none") +
  labs(x = x(),
       y = y())#,
   #tooltip = c("x", "y")
)

  })


#df_filtered <- eventReactive(input$update2, {
#df %>% filter(zipcode %in% input$zipcode,
#             str_type %in% input$str_type,
#             property_type %in% input$property_type)
#})

#renderPlotly({
#p <- ggplot(data = df_filtered()) +
#geom_histogram(aes(x = max_renter_cap, fill = zipcode), binwidth = 1, alpha = 0.5, position = "dodge") +
#labs(title = "Maximum Renter Capacity by Zipcode",
#x = "Maximum Renter Capacity",
#y = "Frequency") +
#scale_fill_discrete(name = "Zipcode")

#ggplotly(p)
#})
```

```{r}
df_boxplot = eventReactive(input$update2, {
  df %>% 
    filter(zipcode %in% input$zipcode2,
             str_type %in% input$str_type2,
             property_type %in% input$property_type2) %>%
    mutate(x = .data[[input$x]],
           y = .data[[input$y]])
})
  
x <- eventReactive(input$update2, input$x)
y <- eventReactive(input$update2, input$y)

renderPlotly({
ggplotly(
  ggplot(data = df_boxplot(), mapping = aes(x = x, y = y)) +
  geom_boxplot() +
  geom_jitter(width=0.2, height=0.1) +
  labs(x = x(),
       y = y())#,
   #tooltip = c("y", "x", "zipcode2", "str_type2")
)

  })

```

## Column

### Information of selected Short-Term Rentals

```{r}

table_colnames <- c("Address", "Zipcode", "STR Type", "Property Type", "Total Bedrooms", "Rented Bedrooms", "Max. Renter Capacity", "Bathrooms", "Kitchen", "Condominium Association")

df_table2 <- eventReactive(input$update2, {
    df %>%
      filter(zipcode %in% input$zipcode2,
             str_type %in% input$str_type2,
             property_type %in% input$property_type2) %>%
      select(address, zipcode, str_type, property_type, total_bedrooms, rented_bedrooms, max_renter_cap, bathrooms, kitchen, condo) %>%
    set_colnames(table_colnames)
  })
  
renderDT({
    datatable(df_table2(), 
              options = list(paging = TRUE,
                             searching = TRUE,
                             orderClasses = TRUE))
  })

```

# About {data-orientation="rows" data-icon="fa-info-circle"}

### **DATA DICTIONARY**

The dataset can be found [here](https://data.cambridgema.gov/Inspectional-Services/Short-Term-Rentals/wxgv-w968/explore/).

The official data dictionary can be found [here](https://data.cambridgema.gov/Inspectional-Services/Short-Term-Rentals/wxgv-w968).

| Column Name                 | Description                                                            |
|---------------------|---------------------------------------------------|
| **ID**                      | Rental ID Number                                                       |
| **Issue Date**              | Date of Issue for the Short Term Rental certificate                    |
| **Address**                 | Address of the STR                                                     |
| **Latitude**                | Latitude of the location                                               |
| **Longitude**               | Longitude of the location                                              |
| **Zipcode**                 | Zipcode of the location                                                |
| **STR Type**                | Response to form question: "Please select the appropriate STR type"    |
| **Property Type**           | Description of the building property type                              |
| **Condominium Association** | Is the Short-Term Rental Unit part of a condominium association?       |
| **Total Bedrooms**          | Total number of legal bedrooms in the short-term rental unit           |
| **Rented Bedrooms**         | Number of legal bedrooms to be rented as part of this application      |
| **Maximum Renter Capacity** | Maximum number of people to which the short term rental will be rented |
| **Kitchen**                 | Will the kitchen be available to the renter?                           |
| **Bathrooms**               | Number of bathrooms available to the renter                            |
| **Airbnb**                  | Whether the rental is listed on this service                           |
| **HomeAway**                | Whether the rental is listed on this service                           |
| **FlipKey**                 | Whether the rental is listed on this service                           |
| **VRBO**                    | Whether the rental is listed on this service                           |
| **Craigslist**              | Whether the rental is listed on this service                           |
| **Couch Surfing**           | Whether the rental is listed on this service                           |
| **Boston Rentals**          | Whether the rental is listed on this service                           |

### **META INFORMATION**

**Title:** Short-term rentals in Cambridge

**Summary:**\
This dataset is part of the **Cambridge's Open Data Program**. It contains information about Short-term rentals (STR) in Cambridge, Massachusetts. STR refers to the rental of any dwelling unit of bedroom as residential accommodation for a duration of less than 30 consecutive days.

Use cases of this dataset include providing valuable insights into the short-term rental market in Cambridge, like understanding the distribution, density and characteristics of STR across Cambridge neighborhoods, which can be useful for a variety of stakeholders, including policymakers, real estate investors, and community organizations.

**Publisher:** [Open Data - City of Cambridge, MA](https://www.cambridgema.gov/departments/opendata)

**Date Created:** April 4, 2018

**Last Updated:** April, 2023

**Data Provided by:** City of Cambridge Inspectional Services Department

**Dataset Owner:** Josh Wolff

**License:** The license for this dataset is unspecified

**Developers:**

-   Joslyn Fu

-   Delina Irgau

-   Michael Sam
